---
title: "Tutorial 7 Worksheet AY 22/23 Sem I"
subtitle: DSA2101
output:
  html_document: default
  pdf_document:
    keep_tex: yes
---

# Good and Bad Graphs

For this section, your tutor will introduce some charts to you during the class,
and discuss their merits and deficiencies. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
l_themes <- read_csv("../data/lego/themes.csv", col_types = cols("i", "c", "i"))
l_sets <- read_csv("../data/lego/sets.csv", 
                     col_types = cols("c", "c", "i", "i", "i"))
l_inv_sets <- read_csv("../data/lego/inventory_sets.csv", 
                     col_types = cols("i", "c", "i"))
l_parts <- read_csv("../data/lego/parts.csv", col_types = cols("c", "c", "i"))
l_part_cat <- read_csv("../data/lego/part_categories.csv",
                     col_types = cols("i", "c"))
l_inv_parts <- read_csv("../data/lego/inventory_parts.csv", 
                     col_types = cols("i", "c", "i", "i", "c"))
l_inv <- read_csv("../data/lego/inventories.csv", 
                     col_types = cols("i", "i", "c"))
l_colors <- read_csv("../data/lego/colors.csv",
                     col_types = cols("i", "c", "c", "c"))
```

# Additional dplyr practice 

Here are some questions for additional practice with dplyr, and data exploration.
The questions use the lego tables from tutorial 6.

1. Theoretically, we can use both the `inventory_parts` table and the `sets` table 
   to compute the number of parts in a set. Do the numbers tally?
   
```{r}
l_inv_parts %>% 
  filter(is_spare == "f") %>% 
  group_by(inventory_id) %>% 
  summarise(num_parts = sum(quantity), .groups="drop") %>% 
  left_join(select(l_inv, 1, 3), by=c("inventory_id" = "id")) %>% 
  left_join(select(l_sets, set_num, num_parts), by="set_num") %>% 
  filter(num_parts.x != num_parts.y)
```

2. Within spare parts, which are the top 5 most common part categories?
```{r}
l_inv_parts %>% filter(is_spare == "t") %>% 
  select(part_num, quantity) %>% group_by(part_num) %>% 
  summarise(qty = sum(quantity), .groups="drop") %>% 
  left_join(select(l_parts, 1, 3), by="part_num") %>% 
  left_join(l_part_cat, ., by=c("id" = "part_cat_id")) %>% 
  slice_max(order_by = qty, n=5)
```

3. Think of one interesting question to ask about this dataset, and try to
   answer it using the data.
   
   Possible ideas:
   1. What is the trend of new sets/number of pieces in each set being released?
   2. What is the trend of unique colours in each set over the years?
   3. Which movie tie-ins have the most sets?
   4. What kinds of sets have the most number of pieces? Is it possible to 
      categorize them by theme?