---
title: "Tutorial 8 Worksheet AY 22/23 Sem 1"
subtitle: DSA2101
output:
  html_document: default
  pdf_document: default
urlcolor: blue
---

The dataset for this tutorial comes from the UCI machine learning repository.
The dataset `wdbc.data` can be obtained from Canvas or from 
[here.](https://archive.ics.uci.edu/ml/datasets/breast+cancer+wisconsin+(diagnostic))

The dataset contains information on 569 cancer patients. We shall use it to 
explore the relationship between certain numerical features - features
computed from images of the cancer cells, and whether the cancer is 
malignant or benign. Hence our response variable is *categorical*, while our 
features are all *numerical*.

Here is a brief description of the columns in the data:

* Col. 1: ID number of patient
* Col. 2: Diagnosis of cancer type: M = malignant and B = benign.
* Col. 3 - 12: mean of ten different real-valued numerical features of cancer cells:
  * radius (mean of distances from center to points on the perimeter)
  * texture (standard deviation of gray-scale values)
  * perimeter
  * area
  * smoothness (local variation in radius lengths)
  * compactness (perimeter^2 / area - 1.0)
  * concavity (severity of concave portions of the contour)
  * concave points (number of concave portions of the contour) 
  * symmetry
  * fractal dimension ("coastline approximation" - 1)
* Col. 13 - 22: resp. standard error of the above features.
* Col. 23 - 32: mean of the largest three of each of the above features. 

We will only be working with the first 12 columns for this tutorial.

## Histogram of radius

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(gridExtra)
```

```{r}
wdbc <- read.csv("../data/wdbc.data", sep = ",", header=FALSE) %>% 
  select(1:12) 
colnames(wdbc) <- c("ID", "diagnosis", "radius", "texture", "perimeter", 
                      "area", "smoothness", "compactness", "concavity", 
                      "concave_points", "symmetry", "fractal_dimension")
```

The following overlays the density plot of radius on the histogram, for each 
type of diagnosis. Recreate the plot as best you can.

```{r fig.align='center', out.width='70%'}
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant")) %>% 
  ggplot(aes(x=radius)) + 
  geom_histogram(aes(y=after_stat(density),fill=diagnosis), 
                 bins=30,
                 position = position_dodge2())+ 
  geom_density(aes(col=diagnosis)) +
  labs(title="Distribution of Radius by Diagnosis", x="Radius", y="Density") + 
  scale_fill_discrete(name="Diagnosis") + scale_color_discrete(name="Diagnosis")
```

Here are some alternative ways in which we can present the two distributions:

1. Simply showing the two densities, differentiated by colour (note that fill 
   aesthetic for densities will fill the area under the density).
```{r}
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant")) %>% 
  ggplot(aes(x=radius)) + 
  geom_density(aes(col=diagnosis))
```

2. Faceting the two distributions and showing both histogram and density geoms.
```{r}
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant")) %>% 
  ggplot(aes(x=radius)) + 
  geom_histogram(aes(y=after_stat(density), fill=diagnosis))  + 
  geom_density(aes(col=diagnosis)) + facet_wrap(~diagnosis, nrow=2)
```

3. Stacking the histograms on top of one another. This is usually the hardest to read/interpret.
```{r}
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant")) %>% 
  ggplot(aes(x=radius)) + geom_histogram(aes(fill=diagnosis))
```

4. To **overlay** the histograms on top of one another, we need to introduce 
   some transparency, and we can achieve the objective in two ways:
   
   First, by manually adding layers one by one. This underlines the point that 
   `ggplot` calls return an object that you can re-use, unlike base R plots.
```{r}
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant")) %>% 
  filter(diagnosis == "Benign") -> wbc1
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant")) %>% 
  filter(diagnosis == "Malignant") -> wbc2

tmp_colours <- scales::hue_pal()(2)
ggplot() + 
  geom_histogram(data=wbc1, aes(x=radius), fill=tmp_colours[1], alpha=0.6) +
  geom_histogram(data=wbc2, aes(x=radius), fill=tmp_colours[2], alpha=0.6)

```

   Alternatively, we can use position identity:
```{r}
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant")) %>% 
  ggplot(aes(x=radius)) + geom_histogram(aes(fill=diagnosis), position="identity",
                                         alpha=0.6)
```

## Proportion of Benign Cancer

The plot above shows the distribution of radius for each type of diagnosis. However,
when our response variable is categorical, we would be more interested in how 
the proportion of benign cancer changes as radius changes. 

```{r}
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant"),
                radius_intervals = cut_interval(radius, length=2)) %>% 
  group_by(radius_intervals) %>% 
  summarise(prop_b =  mean(diagnosis == "Benign"), .groups="drop")  %>% 
  ggplot(aes(x=radius_intervals, y=prop_b)) + geom_point() + 
  geom_line(aes(group=1)) + 
  labs(y="Proportion of Benign cases", x="Radius intervals",
       title="Proportion of Benign Diagnosis falls as Radius increases")
```

How can we update or improve this plot?

* We should reflect the counts within each interval. We could also create 
  intervals that have equal number of observations, but we would have to realign
  the positions at which the x-values are plotted.
* We could show error  bars at each position. 
* If numbers are very unbalanced (which is not in this case), we may want to use 
  a Bayesian method.
* We may also annotate the graph to show where the proportion dips below half.

Here are a couple of such improvements:

1. First, we can introduce error bars to show the uncertainty around the 
   estimate in each bin:
```{r}
wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant"),
                radius_intervals = cut_interval(radius, length=2)) %>% 
  group_by(radius_intervals) %>% 
  summarise(prop_b =  mean(diagnosis == "Benign"), 
            prop_u = prop_b + 2*sqrt(prop_b*(1-prop_b)/n()),
            prop_l = prop_b - 2*sqrt(prop_b*(1-prop_b)/n()),
            .groups="drop")  %>%
  ggplot(aes(x=radius_intervals, y=prop_b)) + geom_point() + 
  geom_line(aes(group=1))  + 
  geom_errorbar(aes(ymin=prop_l, ymax=prop_u), width=0.2)
```

2. Second, we could introduce a second variable, and show how the trend 
   differs conditional on the second variable.
```{r}

wdbc %>% mutate(diagnosis = recode(diagnosis, "B"="Benign", "M"="Malignant"),
                radius_intervals = cut_interval(radius, length=2),
                symmetry_intervals = cut_number(symmetry, n=2)) %>% 
  group_by(radius_intervals, symmetry_intervals) %>% 
  summarise(prop_b =  mean(diagnosis == "Benign"), .groups="drop")  %>% 
  ggplot(aes(x=radius_intervals, y=prop_b,col=symmetry_intervals)) + geom_point() + 
  geom_line(aes(col=symmetry_intervals, group=symmetry_intervals)) 

```



## Data Exploration

Create a plot of your own using radius, diagnosis and one other numerical feature.
How does this update your information of how diagnosis changes with radius?

```{r fig.align='center'}
ggplot(wdbc, aes(x=radius, y=concavity)) + 
  geom_point(aes(col=diagnosis), alpha=0.5)
```

The general impression that as radius is larger, the chance of malignant 
diagnosis increases. However, for small radius as concavity is high, 
there is still a chance for malignant. Similarly, for large radius when concavity 
is low, there is a chance for benign.