---
title: "Tutorial 1 Worksheet AY 22/23 Sem 1"
subtitle: DSA2101
output:
  html_document: default
  pdf_document: default
---

# Birthday paradox

The birthday paradox is commonly introduced in probability classes. It revolves 
around the question:

> In a set of $n$ randomly chosen people, what is the probability that at least 
> two will share a birthday?

Assuming that *all birthdates are equally likely*, it is possible to compute that 
we only require 23 people for this probability to be more than half. In this 
tutorial, we shall relax these assumptions and obtain some answers using simulation.

In all that follows, assume that everyone was born in a non-leap year, and that 
birthdates are coded 1 to 365.

## Classical set up

Here are two possible approaches to the functions. The `table` function tabulates
the count of each value, while `any` is a function that returns TRUE if there is 
at least a single TRUE in a TRUE/FALSE vector.

```{r echo=FALSE}
generate_unif_bdates <- function(n=20) {
  bdates <- sample(1:365, size=n, replace=TRUE)
  ifelse(any(table(bdates) > 1), 1, 0)
}

generate_unif_bdates <- function(n=20) {
  bdates <- sample(1:365, size=n, replace=TRUE)
  ifelse(anyDuplicated(bdates), 1, 0)
} 
```

1. Write an R function `generate_unif_bdates` that 
    * takes in a single argument, `n`, 
    * generates `n` birthdates using `sample`, and finally
    * checks if there is any "clash".
   The function should return TRUE if there *was* a clash, and FALSE otherwise.

```{r}
generate_unif_bdates(20)
```

2. Suppose we were to randomly select $n$ students from NUS, and check to see 
if there were any birthday clashes. Let $X_{i,n}$ be a Bernoulli random variable
that takes on the value 1 if there is a clash and 0 otherwise. Then, if 
all birthdays are equally likely, and are independent of one another, we can compute that
\[
P(X_{i,23} = 1) \approx 0.50
\]
Note that $i$ is simply an indication that this random variable corresponds to 
this particular group of $n$ individuals. 

To estimate this probability using simulation, we would generate multiple observations
of $X_{i,23}$ and compute the proportion of times the value 1 (or TRUE) occurs.

`generate_unif_bdates(23)` generates a single observation of $X_{i,23}$. Repeat
this 100 times to see if you obtain a number close to 0.50.

```{r}
X_out <- replicate(100, generate_unif_bdates(23))
mean(X_out)

# An alternative approach: write a wrapper function that allows you to 
# modify the number of simulations as well:
# This is in preparation for the next task.
wrapper_fn <- function(n, nsim) {
  mean(replicate(nsim, generate_unif_bdates(n)))
}
```

3. Now run 100 simulations each for `n` from 10 to 50. Create a similar graph to 
this one:

```{r fig.align='center'}
# Set seed here?
output_prop <- sapply(10:50, 
                      function(x) mean(replicate(100, generate_unif_bdates(x))))

# Using the wrapper function from above, the sapply call is cleaner:
output_prop <- sapply(10:50, wrapper_fn, nsim=100) 

plot(10:50, output_prop, type="b", ylim=c(0,1), xlab="No. of individuals in a group",
     ylab="Prob.", main="Probability of birthday clash")
abline(h=0.5, lty=2, col="red")
abline(v=23, lty=2, col="black")
```

4. As you observed, we do not all get exactly the same graph. We can compute 
Confidence Intervals to quantify this uncertainty. To compute 95\% confidence 
intervals for the probabilities above, we can use this formula:
\[
\hat{p} \;\pm\; z_{0.975}\sqrt{\frac{\hat{p}(1- \hat{p})}{100}}
\]
  where:
  * $z_{0.975}$ is the .975-quantile of the N(0,1) distribution. (`qnorm()` function in R)
  * $\hat{p} = \frac{1}{100}\sum_{i=1}^{100} X_{i, n}$
  
  Apply the formula above to update your plot:
```{r fig.align='center'}
upper <- output_prop + qnorm(0.975)*sqrt(output_prop*(1-output_prop)/100)
lower <- output_prop - qnorm(0.975)*sqrt(output_prop*(1-output_prop)/100)

plot(10:50, output_prop, type="b", ylim=c(0,1), xlab="No. of individuals in a group",
     ylab="Prob.", main="Probability of birthday clash")

lines(10:50, upper, col="red", lty=2)
lines(10:50, lower, col="red", lty=2)
segments(10:50, upper, y1=lower, lend=0)

# An alternative, pre-packaged function for error bars:
#Hmisc::errbar(10:50, output_prop, upper, lower)
```
  
6. How many people do we need if we wish for the probability of 
    * 3 birthdays on one day to be more than 0.75?
    * birthdays within 7 days to be more than 0.5?
    
```{r}
generate_unif_bdates2 <- function(n=20) {
  bdates <- sample(1:365, size=n, replace=TRUE)
  any(table(bdates) > 2) 
}
output_prop2 <- sapply(10:150, 
                      function(x) mean(replicate(100, generate_unif_bdates2(x))))

plot(10:150, output_prop2, type="b", ylim=c(0,1), xlab="No. of individuals in a group",
     ylab="Prob.", main="Probability of birthday clash")
```

```{r}
generate_unif_bdates3 <- function(n=20) {
  bdates <- sort(sample(1:365, size=n, replace=TRUE))
  
  crossover <- c(bdates[bdates >= 359], bdates[bdates<=7]+365)
  if(length(crossover) >=2){
    if(any(diff(crossover)<=7))
      return(TRUE)
  }
  any(diff(bdates) <= 7)
}
output_prop3 <- sapply(2:30, 
                      function(x) mean(replicate(100, generate_unif_bdates3(x))))

plot(2:30, output_prop3, type="b", ylim=c(0,1), xlab="No. of individuals in a group",
     ylab="Prob.", main="Probability of birthday clash")
```


# References

1. [Wikipedia entry on Birthday paradox](https://en.wikipedia.org/wiki/Birthday_problem)