---
title: "Tutorial 5 Worksheet AY 22/23 Sem 1"
subtitle: DSA2101
output:
  html_document: default
  pdf_document:
    keep_tex: yes
---

# Restaurants Data

In chapter 2, we worked with the JSON version of the restaurants data. The 
following code converts the file into a tibble within R. Let us use it to 
practice with `dplyr` data manipulations.

```{r message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(jsonlite)

rest_json <- stream_in(file("../data/restaurants_dataset.json")) %>% 
  unnest(cols=c(address, grades))
```

Recall that in the lecture, there 25359 restaurants. The above procedure has 
removed the 738 records with NULL in the grades section, but it retains those with 
"Not Yet Graded".

The output tibble should look like this:
```{r echo=FALSE}
head(rest_json)
```

1. As you can see, the `coord` column contains a list instead of numeric elements. 
  Separate `coord` into two columns, lat and long, and drop the original `coord` column.
  Assign the name `q1_tbl` to the new tibble. It should look like this:
```{r }
q1_tbl <- mutate(rest_json, 
                   lat = sapply(coord, `[`, 2), long=sapply(coord, `[`, 1), 
                   coord = NULL, .after=building)

# mutate can also work with a function that returns multiple columns:
# 
get_lat_long <- function(list_vec) {
  lat <- sapply(list_vec, `[`, 2)
  long <- sapply(list_vec,  `[`, 1)
  data.frame(lat = lat, long = long)
}
q1_tbl <- mutate(rest_json, 
                 get_lat_long(coord), coord=NULL, .after=building)
```

2. Find the range of lat and long for the restaurants in `q1_tbl`. Are there 
   mistakes in the dataset? Identify one restaurant with incorrect coordinates.
```{r}
q1_tbl %>% select(lat, long) %>% summary()
# filter(q1_df, lat < 40 | lat > 41 | long < -75 | long > -73) %>% View()
```

2. How many restaurants in Manhattan serve American cuisine?

If you inspect the cuisine types, you would find that American is listed with
an extra space at the back. We can trim off that space first before listing the
count.

```{r}
q1_tbl %>% mutate(cuisine = str_trim(cuisine)) %>%
  filter(borough == "Manhattan", cuisine == "American") %>%
  select(restaurant_id) %>% unique() %>% NROW
# 3199 restaurants
```

4. Find the restaurants that have been graded the most number of times. 
   Include the restaurant name in your output. Store your output as `q4_tbl`. 
   The columns should match this output (only first 2 rows shown):
```{r}
q4_tbl <- q1_tbl %>% filter(grade != "Not Yet Graded") %>%
  group_by(restaurant_id) %>%
  summarise(name = name[1], n=n(), .groups="drop") %>%
  slice_max(n, n = 1)
head(q4_tbl, n=2)
```

5. For each restaurant, compute the shortest duration (in days) between gradings. 
   You may want to take a look at the `difftime` function in R.
```{r}
q5_tbl <- q1_tbl  %>% #head() %>%
  filter(grade != "Not Yet Graded") %>%
  arrange(restaurant_id, desc(date)) %>%
  mutate(d2 = as.numeric(difftime(lag(date), date, units="days"))) %>%
  group_by(restaurant_id) %>%
  filter(row_number() > 1) %>%
  filter(d2 == min(d2)) %>% ungroup() %>% 
  select(restaurant_id, name, d2) %>% arrange(d2)
head(q5_tbl)
```

6. The 5-number summary of a dataset consists of the minimum, 1st-quartile,
   median, 3rd-quartile and maximum of the dataset. Compute the five-number summary
   score for each grade. Include the count for each grade.
```{r}
fivenum <- function(x) {
  out <- as.vector(summary(x))
  data.frame(min=out[1], q1=out[2], median=out[3], q3=out[5], max=out[6])
}
q6_tbl <- q1_tbl %>% filter(grade != "Not Yet Graded") %>%
  group_by(grade) %>%
  summarise(fivenum(score))
```

7. Count the number of gradings in each calendar month in each borough.  It is
   ok for month to be of class "character".
```{r}
q7_tbl <- q1_tbl %>% mutate(month = cut(date, breaks="month"),
                    yr = cut(date, breaks="year")) %>%
  group_by(borough, yr, month) %>%
  summarise(month_total=n(), .groups = "drop_last") %>%
  mutate(yr_total = sum(month_total)) %>% ungroup() %>%
  select(-yr)
q7_tbl
```

8. Recode the following cuisines as `Asian`, and then find the proportion of Asian
   restaurants in each borough:
      * Vietnamese/Cambodian/Malaysia
      * Thai
      * Chinese
      * Chinese/Japanese
      * Pakistani
      * Korean
      * Indonesian
      * Indian
      * Asian
   You may want to take a look at `recode()`.

```{r}
q8_tbl <- q1_tbl %>% mutate(cuisine2 = recode(cuisine,
                                      `Vietnamese/Cambodian/Malaysia`="Asian",
                                      Thai = "Asian", Chinese = "Asian",
                                      `Chinese/Japanese` = "Asian",
                                      Pakistani = "Asian",
                                      Korean = "Asian", Indonesian = "Asian",
                                      Indian = "Asian",
                                      default = cuisine)) %>%
  group_by(borough) %>%
  summarise(prop = sum(cuisine2 == "Asian")/n(), .groups="drop")
```

9. Use `across` to compute the min. and max. letter grade and score for each
   restaurant.
```{r}
q1_tbl %>% filter(grade != "Not Yet Graded") %>%
  group_by(restaurant_id, name ) %>%
  summarise(across(c(grade, score), .fns=list(min = min, max=max)),
            num_gradings=n())
```

10. List the names of all restaurants that have "C" as their first grade, and
   "A" as the second grade.

```{r}
q10_tbl <- q1_tbl %>% group_by(restaurant_id) %>%
  summarise(tf = if_else((first(grade, date) == "C") &&
                           (nth(grade, 2, date) == "A"), 1, 0),
            name = name[1]) %>%
  filter(tf == 1)  %>% sample_n(15)

q10a_tbl <- q1_tbl %>% group_by(restaurant_id) %>%
  arrange(date, .by_group = TRUE) %>%
  summarise(tf = if_else(grade[1] == "C" && grade[2] == "A", 1, 0),
            name = name[1]) %>%
  filter(tf == 1)  %>% sample_n(15)
```

11. List all restaurants with only "A" grades.

```{r}
q11_tbl <- q1_tbl %>%
  group_by(restaurant_id) %>%
  filter(!(any(grade != "A")))%>%
  ungroup() %>% 
  select(restaurant_id,name) %>% unique()
```