---
title: "Assignment 01"
author: "Claudeon Reinard Susanto"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

```{r data}
data <- readRDS("../data/assignment01.rds")
akima_data <- data$akima_data
artist_data <- data$artist_data
```
# Akima Interpolation
## Question 1
`compute_slopes` will return the estimated slope at `x[3]` given two vectors of length 5
```{r compute-slopes}
compute_slopes <- function(x, y) {
  # condition: both x and y are vectors of length 5
  m <- sapply(1:4, function(index) (y[index+1]-y[index])/(x[index+1]-x[index]))
  d <- c(abs(m[2]-m[1]), abs(m[4]-m[3]))
  s <- ifelse((d[1]==0 && d[2]==0), 
              1/2*(m[2]+m[3]),
              (d[2]*m[2]+d[1]*m[3])/(mean(d)))
  s
}

```

## Question 2
Finding interpolated points between two given points
```{r est-poly}
est_poly <- function(x, y, slopes, n) {
  deltaY <- y[2]-y[1]
  deltaX <- x[2]-x[1]
  p0 <- y[1]
  p1 <- slopes[1]
  p2 <- (3*deltaY/deltaX - 2*slopes[1]-slopes[2])/deltaX
  p3 <- (sum(slopes)-2*deltaY/deltaX)/(deltaX^2)
  
  poly <- function(v) {
    newV <- v-x[1]
    return(p0 + p1*newV+p2*newV^2+p3*newV^3)
  }
  
  x_vals <- seq(x[1],x[2],length.out=n)
  interpolated <- poly(x_vals)
  ret <- cbind(x_vals, interpolated)
  ret
}
```

## Question 3
```{r akima-plot, fig.align='center', fig.cap="Akima Interpolation"}
x <- akima_data[,1]; y <- akima_data[,2];
n <- length(x)
x <- sort(x); y <- sapply(1:n, function(i) akima_data[which(akima_data[,1]==x[i]),2])

# compute slope for x[3], x[4], ... x[n-2]
slopes <- sapply(seq(3,n-2,by=1), function(i) 
  compute_slopes(x[(i-2):(i+2)], y[(i-2):(i+2)]))

pts <- do.call(rbind, lapply(seq(3,n-3,by=1), function(i) est_poly(c(x[i],x[i+1]), c(y[i],y[i+1]), slopes[(i-2):(i-1)],n=5)))

plot(x,y,col="red",pch=19)
lines(pts,col="blue")

```

## Question 4
First we will clean up `gender`.
```{r gender-unique, collapse=TRUE}
unique(artist_data$gender)
```
Based on this result, we can convert all string containing `"F"` or `"f"` to `Female`

```{r}
gender_clean <- as.factor(str_replace(artist_data$gender, regex("F[\\S ]+$", ignore_case=TRUE), "Female"))
```
Next we will clean up `dates`
```{r}
# "XXXX-XXXX" format
pattern1 <- "([0-9]{4})\\S([0-9]{4})"
# "born XXXX" format
pattern2 <- "^born +(\\d{4})$"
# "XXXX/XX-XXXX" format
pattern3 <- "\\d{4}/\\d+\\S\\d{4}"
# "b. XXXX" format
pattern4 <- "b. \\d{4}"
# "XXXX/XX" format
pattern5 <- "\\d{4}/\\d+"

dates_clean <- artist_data$dates
dates_clean <- str_replace(dates_clean, "([0-9]{4}) ?\\S ?([0-9]{4})", "\\1-\\2")
dates_clean <- str_replace(dates_clean, "^[Bb]orn +(\\d{4})$", "born \\1")
dates_clean <- str_replace(dates_clean, "(\\d{4}) ?or ?(\\d+) ?\\S ?(\\d{4})", "\\1/\\2-\\3")
dates_clean <- str_replace(dates_clean, "[Bb]. (\\d{4})", "b. \\1")
dates_clean <- str_replace(dates_clean, "(\\d{4}) ?- ?(\\d{2})[$ ]", "\\1/\\2")
```
The result is stored in `artist_data_clean`
```{r, collapse=TRUE}
artist_data_clean <- artist_data
artist_data_clean$gender <- gender_clean
artist_data_clean$dates <- dates_clean
head(artist_data_clean)
```
## Question 5
```{r artist-birth-years, collapse=TRUE}
artist_birth_years <- sort(as.numeric(unique(unlist(str_extract_all(artist_data_clean$dates, "(?<=born )(\\d{4})")))))
artist_birth_years
```

## Question 6
First, we identify those with the format `<city>, <country>`
```{r artist-birth-country, collapse=TRUE}
format_1commas <- sapply(lapply(str_split(artist_data$placeOfBirth, ","),str_trim), length) == 2
countries_1commas <- sapply(str_split(artist_data[format_1commas,]$placeOfBirth, ","), 
                    function(x) {
                      x <- str_trim(x);
                      return(x[2]);
                      })
countries_1commas <- unique(countries_1commas)
countries_1commas
```
We are left with the entries whose formatting is different from the more common one (i.e. with zero or two commas).
We start with dealing with those containing only one word (zero commas)
```{r, collapse=TRUE}
format_0commas <- sapply(lapply(str_split(artist_data$placeOfBirth, ","),str_trim), length) == 1
countries_0commas <- artist_data$placeOfBirth[format_0commas]
na_entries <- countries_0commas==""
countries_0commas <- unique(countries_0commas[!na_entries])
countries_0commas[!(countries_0commas %in% countries_1commas)]
```
We can select the entries which are clearly names of countries
```{r}
countries_0commas <- c("Barbados", "Singapore", "Armenia", "Myanmar", "Zimbabwe", "Chung-hua Min-kuo", "Lietuva", "Costa Rica", "Nigeria", "Samoa")
```
Moving on for those with 2 commas
```{r, collapse=TRUE}
format_2commas = !(format_1commas | format_0commas)
countries_2commas <- artist_data$placeOfBirth[format_2commas]
countries_2commas
```
Therefore we can create `countries_2commas`
```{r, collapse=TRUE}
countries_2commas <- c("France", "Colombia", "Yisra'el", "Hrvatska")
```
The resulting `artist_birth_country` is
```{r, collapse=TRUE}
artist_birth_country <- unique(c(countries_0commas, countries_1commas, countries_2commas))
sort(artist_birth_country)
```