---
title: "Practice I Example Solutions"
author: "Vik"
output:
  html_document: default
  pdf_document: default
---
```{r message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(classInt)
```

# Question 1: Monthly Revenue

## Question Statement

The monthly revenue is computed as `sum(UnitPrice * Quantity)`. Comment on the
quality of the following plot, which is meant to depict the monthly revenue that
the company receives. Is it accurate? *Hint: Check the range of invoice dates.*

If you feel that there are mistakes in this plot, produce a corrected one. On
the other hand, if you feel this plot cannot be improved, create a variation of
it using at least one different geom or aesthetic.

```{r echo=FALSE, fig.align="center", out.width="70%"}
retail_clean <- readRDS("../data/retail_clean.rds") %>% 
  filter(UnitPrice > 0) %>% 
  mutate(idate = as.Date(InvoiceDate))
  
monthly_rev <- mutate(retail_clean, month = month(InvoiceDate, label=FALSE)) %>% 
  group_by(month) %>% 
  summarise(rev = sum(UnitPrice*Quantity)/1e3, .groups="drop")
ggplot(monthly_rev) +
  geom_col(aes(x=factor(month, levels=1:12), y=rev)) +
  labs(title="Monthly Revenue", y="Thousand Pounds", x="Month")
```

## Solution

Here is a corrected version of the plot. It separates the two December months,
which have only been partially observed.

```{r fig.align="center", out.width="75%"}
monthly_rev <- mutate(retail_clean, yr_mth = strftime(InvoiceDate, "%Y-%m")) %>%
  group_by(yr_mth) %>%
  summarise(rev = sum(UnitPrice*Quantity)/1e3,
            days = as.numeric(max(InvoiceDate) - min(InvoiceDate)),
            .groups="drop") %>%
  mutate(project = c(1, rep(0, 11), 1)) %>%
  mutate(projected = if_else(project==1, rev/days * 31, rev))
ggplot(monthly_rev) +
  geom_col(aes(x=factor(yr_mth), y=projected, fill=as.factor(project)),
           show.legend = FALSE) +
  labs(title="Monthly Revenue", y="Thousand Pounds", x="Month",
       subtitle="Revenue for Dec. months have been estimated from fewer than 31 days.")
```

# Question 2: Christmas Shopping

## Question statement

The company would like to figure out when people begin Christmas shopping, so
that they can start preparing for it. In the box below, write the R code that
will perform the following tasks:

In this question, a line item refers to a record in the dataset. For instance, 
the first line item refers to StockCode 85123A bought in Invoice 536365.

1. Create a new column from `InvoiceDate`, that is of class `Date`.
2. Use a regular expression to count the number of *line items* each day, that 
are related to Christmas. Look for the following keywords in the Description of 
each item: CHRISTMAS, XMAS, SANTA or REINDEER.
3. Create a visualisation of this daily count, and use it to conclude when 
Christmas shopping begins.

## Solution

```{r fig.align="center"}
xmas <- filter(retail_clean,
               str_detect(Description,
                          "(CHRISTMAS)|(XMAS)|(SANTA)|(REINDEER)")) %>%
  group_by(idate) %>% count() %>% ungroup() 
ggplot(xmas) + geom_line(aes(x=idate, y=n)) +
  labs(x='Date', y='Number of item orders',
       title='How early does Xmas shopping begin?')
```

Let us compute the mean number of detections of Xmas items for each month, along 
with the standard deviation.

```{r}
xmas %>% mutate(mths= cut(idate, "months")) %>% 
  group_by(mths) %>% 
  summarise(mm = mean(n), sd=sd(n), .groups="drop") %>% 
  knitr::kable()
```

It looks as though the month of July has a dramatic pick-up from June. Hence we
conclude Xmas shopping begins in July.

# Growth outside United Kingdom

## Question statement

The company in question is located within the United Kingdom (UK). Naturally,
most of the revenue comes from within the UK. In this question, we shall 
investigate the change in revenue to countries in Asia.

Use the country table `cty_coords.rds` to create the following chart, where
point size indicates the revenue from Asian countries in the *third* quarter of
2011.

```{r}
cty <- readRDS("../data/cty_coords.rds")
out_UK <- left_join(retail_clean, cty, by=c("Country"="rtl")) %>% 
  mutate(quarter = cut(InvoiceDate, "quarters")) %>% 
  filter(quarter == "2011-07-01", group == "Asia") %>% 
  group_by(Country) %>% 
  summarise(total_rev = sum(UnitPrice * Quantity),
            long=long[1], 
            lat = lat[1], .groups="drop")
out_UK$label <- c("Australia, 32K", "Israel, 6K", "Japan, 4K", 
                  "Singapore, 4K", "UAE, 1K")

ggplot(out_UK) + geom_point(aes(x=long, y=lat, size = total_rev),
                            alpha=0.8, colour ="royalblue2") + 
  coord_fixed() + 
  geom_text(aes(x=long, y=lat, label=label, 
                hjust=c(1.2, -.2, 1.1, -.1, -.2)), size=3) +
  labs(title="Revenue from Asian countries, Q3 2011", 
       y="Latitude", x="Longitude") + 
  scale_size(name="Revenue")
```

# Customer features

## Question statement

The table `cust_features.rds` contains the following features for each customer 
id in the retail table:

1. Country: The country in which the customer is located.
2. CustomerID: The customer id.
3. m_amt: The average revenue per invoice to this customer.
4. m_item: The average number of items ordered by this customer, per invoice.
5. m_uq: The average number of unique items ordered by this customer, per invoice.
6. n_inv: The number of invoices issued to this customer.
7. group: A grouping of countries into continents, except for United Kingdom.
8. grp2: Same as `group`, except for an abbreviation of United Kingdom to UK.

Here is one plot that begins to explore the relationship between unique items 
bought and total number of items bought by a customer.

```{r echo=FALSE}
cust_features <- group_by(retail_clean, Country, CustomerID, InvoiceNo) %>%
  summarise(amt = sum(UnitPrice * Quantity),
            item_total = sum(Quantity),
            item_uq = n_distinct(StockCode), .groups="drop_last") %>%
  summarise(m_amt = mean(amt),
            m_item = mean(item_total),
            m_uq = mean(item_uq),
            n_inv = n_distinct(InvoiceNo))
cust_features <- left_join(cust_features,
                           select(cty, rtl, group),
                           by=c("Country" = "rtl")) %>%
  ungroup()
cust_features <- mutate(cust_features,
                        grp2 = if_else(Country == "United Kingdom", "UK", group))
```

```{r}
#filter(cust_features, m_uq > 1, grp2 != "S_America", !is.na(grp2)) %>%
#  ggplot() +
#  geom_point(aes(x=m_item, y=m_uq, col=grp2), alpha=0.2, show.legend = FALSE) +
#  facet_wrap( ~ grp2) +
#  scale_x_log10() + scale_y_log10()

filter(cust_features, m_uq > 1, grp2 != "S_America", grp2 != "NA", grp2!='Africa') %>%
  ggplot() +
  geom_point(aes(x=m_item, y=m_uq, col=grp2), alpha=0.3, show.legend = FALSE) +
  facet_wrap( ~ grp2) +
  scale_x_log10() + scale_y_log10() + 
  labs(title="Number of unique items and total number of items", 
       x="Total number of items", y="Unique",
       subtitle="(103 customers with only one unique item removed)")
```


