---
title: "Practice III soln"
output:
  pdf_document: default
  html_document: default
---

# Questions

The dataset in `Real estate valuation data set.xlsx` contains real estate prices
in Sindian (Xindian) District of New Taipei City in Taiwan. The seven inputs are
as follows:

| Column name   | Description                                                |
| ------------- | :--------------------------------------------------------- |
| X1            | transaction date. (for example, 2013.250=2013 March, etc.) |
| X2            | age of the house. (unit: years)                            |
| X3            | distance to the nearest MRT station. (unit: meters)        |
| X4            | number of convenience stores within walking distance.      |
| X5            | latitude. (unit: degree)                                   |
| X6            | longitude. (unit: degree)                                  |

The output variable, price, is as follows:

*  Y = house price of unit area (10000 New Taiwan Dollar/Ping, where Ping is a 
local unit, 1 Ping = 3.3 meter squared) 


## Question 1 (15 marks)

Re-create the following plot as closely as you can. You may need the following 
information:

1. The EPSG code 3825 was used to project the coordinates.
2. If the hex bins are not available to you, use `stat_summary_2d`.
3. The fill aesthetic was mapped the the median house price in each bin.

```{r fig.align='center', out.width='70%', message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(readxl)

re <- read_excel("../data/Real estate valuation data set.xlsx")
colnames(re) <- c("id", "trans_date", "house_age", "dist_MRT", "num_stores", 
                  "lat", "long", "price")

re_sf <- st_as_sf(re, coords = c("long", "lat"), crs="+proj=longlat +datum=WGS84")
re_sf2 <- st_transform(re_sf, 3825) # convert to units of metres.
re_coords <- st_coordinates(re_sf2)
re2 <- bind_cols(re, data.frame(re_coords)) 

ggplot(re2, aes(x=X, y=Y, z=price)) + 
  scale_fill_binned(breaks=c(27,38,46), type="viridis") + 
  stat_summary_hex(fun="median")  + coord_fixed() + 
  labs(title = "Binned median property prices", fill="Median price", x="", y="") + 
  theme(legend.position = "bottom")
```

## Question 2 (15 marks)

A property analyst informs you that the centre of town is at (506800, 2765500).
Add a column to your dataset that contains the Euclidean distance the the centre
of town. 

Now, create **two** graphics that demonstrate how price varies with the other variables. 
Please ensure that the following criteria are satisfied:
    * Distance to the centre of town must be present in at least one of the graphs.
    * The price variable must be present in both of them.

```{r fig.align='center', out.width='70%'}
X_centre <- 506800
Y_centre <- 2765500

re2 <- mutate(re2, dist_centre = sqrt((X - X_centre)^2 + ( Y - Y_centre)^2))

ggplot(re2, aes(x=dist_centre, y=price)) + geom_point() + 
  geom_smooth(method="lm") + facet_wrap(~num_stores) + scale_x_sqrt() +
  labs(title="Distance to centre and number of stores")
```

## Question 3 (10 marks)

The same property analyst suspects that price has a non-linear relationship 
with house age: For modern houses, the price decreases with age but for 
vintage houses, the prices increases with age.

The problem is that the change-point is unknown - at what age should we classify 
a house as vintage? The following function will return the parameters for a line 
of best fit through these points, given a change-point. 

```{r}
get_gradient_intercept <- function(price, age, breakpoint) {
  modern <- pmin(age - breakpoint, 0)
  vintage <- pmax(age - breakpoint, 0)
  lm0 <- lm(price ~ modern + vintage)
  coef(lm0)
}
```

For instance, *if the analyst believed* the change-point was at 10 years:

```{r}
(coefs <- get_gradient_intercept(re2$price, re2$house_age, 10))
```

Then the fitted line would be plotted like this:

```{r echo=FALSE, fig.align='center', out.width='70%'}
xvals1 <- 1:10
yvals1 <- coefs[1] + coefs[2]*(xvals1 - 10)
xvals2 <- 10:45
yvals2 <- coefs[1] + coefs[3]*(xvals2 - 10)

df1 <- data.frame(x= c(xvals1, xvals2), y= c(yvals1, yvals2))
ggplot(re2) + geom_point(aes(x=house_age, y=price), alpha=0.2) + 
  geom_line(data=df1, aes(x=x, y=y), col="red") + 
  annotate("label", x=c(7.5, 10, 30), y=c(55, 30, 30), 
           label=c("slope = -1.88", "y-value = 34.36", "slope = 0.09"))
```

Find the optimal line of best fit. Desribe what criteria you used to find it, 
and create a properly annotated plot similar to the one above that summarises
the relationship.

One way to do this is to vary the changepoint and compute the one that minimises 
the sum of the squared residuals.

```{r}
get_gradient_intercept2 <- function(price, age, breakpoint) {
  modern <- pmin(age - breakpoint, 0)
  vintage <- pmax(age - breakpoint, 0)
  lm0 <- lm(price ~ modern + vintage)
  mean(residuals(lm0)^2)
}
rmses <- sapply(1:43, get_gradient_intercept2, price=re2$price, age=re2$house_age)
which.min(rmses)
```


```{r echo=FALSE, fig.align='center', out.width='70%'}
coefs2 <- get_gradient_intercept(re2$price, re2$house_age, 24)
xvals1 <- 1:24
yvals1 <- coefs2[1] + coefs2[2]*(xvals1 - 24)
xvals2 <- 24:45
yvals2 <- coefs2[1] + coefs2[3]*(xvals2 - 24)

df2 <- data.frame(x= c(xvals1, xvals2), y= c(yvals1, yvals2))
ggplot(re2) + geom_point(aes(x=house_age, y=price), alpha=0.2) + 
  geom_line(data=df2, aes(x=x, y=y), col="red")  + 
  labs(title="Houses can be classifed as vintage after 24 years", 
       x="House age (years)", y="Price")
```